---
title: "Normalize data"
author: "TB"
date: "11/21/2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("magrittr")
library("openxlsx")
```

This notebook creates the files analysis.xlsx, sector_analysis_per_experiment.xlsx, 


```{r}
csv_tracks <- 'tracks.csv'
csv_normalized <- 'normalized.csv'
analysis_folder <- '../data/'
results_folder <- '../results/'

# grouping variables 
strata <- c('Metadata_condition','Metadata_dose','Metadata_matrix')


```

# load data 

```{r} 
normalized <- read_csv(file.path(analysis_folder, csv_normalized)) %>%
  filter(Track_Length > 19)
```
Summarize and export mean and sem values for each EXPERIMENT
```{r}
analysis <- normalized %>% group_by_(.dots = strata) %>%
  summarise(
    "mean directionality" = mean(Track_Directionality),
    "sem directionality" = sd(Track_Directionality) / sqrt(n()),
    "mean speed" = mean(Track_Speed),
    "sem speed" = sd(Track_Speed) / sqrt(n()),
    "mean xFMI" = mean(Track_xFMI),
    "sem xFMI" = sd(Track_xFMI) / sqrt(n()),
    "mean yFMI" = mean(Track_yFMI),
    "sem yFMI" = sd(Track_yFMI) / sqrt(n()),
    n = n()
  ) %>%
  select(Metadata_matrix, Metadata_condition, Metadata_dose, everything()) %>% 
  arrange(Metadata_matrix, Metadata_condition)
```

Export data as xls and csv
```{r}
write_csv(analysis, file.path(results_folder, "analysis.csv"))
write.xlsx(analysis, file.path(results_folder, "analysis.xlsx"))
```



###############################################################################
# Summarize and export mean and sem values for each SECTOR of each EXPERIMENT

```{r}
sector_analysis_temp <- normalized %>% 
  group_by_(.dots = c(strata,"Track_Sector", "Metadata_id")) %>%
  summarise(
    "mean directionality" = mean(Track_Directionality),
    "sem directionality" = sd(Track_Directionality) / sqrt(n()),
    "mean speed" = mean(Track_Speed),
    "sem speed" = sd(Track_Speed) / sqrt(n()),
    "mean xFMI" = mean(Track_xFMI),
    "sem xFMI" = sd(Track_xFMI) / sqrt(n()),
    "mean yFMI" = mean(Track_yFMI),
    "sem yFMI" = sd(Track_yFMI) / sqrt(n()),
    n = n()
  ) %>%
  select(Metadata_matrix, Metadata_id, Metadata_condition, Metadata_dose, everything()) %>% 
  arrange(Metadata_matrix, Metadata_condition, Metadata_id) %>%
  print
```

Additionaly the fraction of tracks in each sector is summarized
```{r}
n_per_experiment <- normalized %>% 
  group_by_(.dots = c(strata, "Metadata_id")) %>%
  summarise("n_per_experiment" = n()) %>%
  print

sector_analysis <- left_join(
  sector_analysis_temp, 
  n_per_experiment,
  by = c("Metadata_condition", "Metadata_dose", "Metadata_matrix", "Metadata_id")
  ) %>%
  mutate("fraction per sector in %" = n / n_per_experiment) %>% 
  print
```

Export data as xls and csv
```{r}
write_csv(sector_analysis, file.path(results_folder, "sector_analysis_per_experiment.csv"))
write.xlsx(sector_analysis, file.path(results_folder, "sector_analysis_per_experiment.xlsx"))
```

Sector analysis per condition
```{r}
sector_analysis_per_condition <- normalized %>% 
  group_by_(.dots = c(strata,"Track_Sector")) %>%
  summarise(
    "mean directionality" = mean(Track_Directionality),
    "sem directionality" = sd(Track_Directionality) / sqrt(n()),
    "mean speed" = mean(Track_Speed),
    "sem speed" = sd(Track_Speed) / sqrt(n()),
    "mean xFMI" = mean(Track_xFMI),
    "sem xFMI" = sd(Track_xFMI) / sqrt(n()),
    "mean yFMI" = mean(Track_yFMI),
    "sem yFMI" = sd(Track_yFMI) / sqrt(n()),
    n = n()
  ) %>%
  select(Metadata_matrix, Metadata_condition, Metadata_dose, everything()) %>% 
  arrange(Metadata_matrix, Metadata_condition) %>%
  print
```

Export data as xls and csv
```{r}
write_csv(sector_analysis_per_condition, file.path(results_folder, "sector_analysis.csv"))
write.xlsx(sector_analysis_per_condition, file.path(results_folder, "sector_analysis.xlsx"))
```

###############################################################################
# Export subpopualtion walker / sneaker / runner or FN and HEM separetly 

```{r}
quantiles <- normalized %>% 
  filter(Metadata_condition == "vehicle") %>%
  filter(Metadata_matrix == "FN") %>%
  extract2("Track_Speed") %>%
  quantile(.,probs = c(.25, .75)) %>% 
  as.vector()

movements <- normalized %>% 
  filter(Metadata_matrix == "FN",Metadata_dose != 0.5, Metadata_dose != "control") %>%
  mutate(sneaker = (Track_Speed < quantiles[[1]])) %>%
  mutate(walker = ( (Track_Speed > quantiles[[1]] )  & ( Track_Speed < quantiles[[2]] ) ) )  %>%
  mutate(runner = (Track_Speed > quantiles[[2]])) %>%
  print

subpopulation_speed <- movements %>%  
  ungroup() %>% 
  group_by(Metadata_dose, Metadata_condition) %>%
  summarise("mean sneaker" = mean(sneaker), "sem sneaker" = sd(sneaker)/sqrt(n()),
            "mean walker" = mean(walker), "sem walker" = sd(walker)/sqrt(n()),
            "mean runner" = mean(runner), "sem runner" = sd(runner)/sqrt(n()), 
            n = n()
    ) %>% print

subpopulation_speed %>% 
  write_csv(file.path(results_folder, "subpopulation_speed_fn.csv"))

subpopulation_speed %>% 
  write.xlsx(file.path(results_folder, "subpopulation_speed_fn.xlsx"))

```

```{r}
quantiles <- normalized %>% 
  filter(Metadata_condition == "vehicle") %>%
  filter(Metadata_matrix == "HEM") %>%
  extract2("Track_Speed") %>%
  quantile(.,probs = c(.25, .75)) %>% 
  as.vector()

movements <- normalized %>% 
  filter(Metadata_matrix == "HEM",Metadata_dose != 0.5, Metadata_dose != "control") %>%
  mutate(sneaker = (Track_Speed < quantiles[[1]])) %>%
  mutate(walker = ( (Track_Speed > quantiles[[1]] )  & ( Track_Speed < quantiles[[2]] ) ) )  %>%
  mutate(runner = (Track_Speed > quantiles[[2]])) %>%
  print

subpopulation_speed <- movements %>%  
  ungroup() %>% 
  group_by(Metadata_dose, Metadata_condition) %>%
  summarise("mean sneaker" = mean(sneaker), "sem sneaker" = sd(sneaker)/sqrt(n()),
            "mean walker" = mean(walker), "sem walker" = sd(walker)/sqrt(n()),
            "mean runner" = mean(runner), "sem runner" = sd(runner)/sqrt(n()), 
            n = n()
    ) %>% print

subpopulation_speed %>% 
  write_csv(file.path(results_folder, "subpopulation_speed_hem.csv"))

subpopulation_speed %>% 
  write.xlsx(file.path(results_folder, "subpopulation_speed_hem.xlsx"))

```
